Improving the Accuracy of a Sequential Mining Model for Hurricane Trajectory Prediction

### 2018-01-29 paper/manuscript_revised_jan25.tex
minor spelling fixes

There are two categories of motivation behind this work. The first motivation of this work is to add to the total investigation completed in the data mining-based hurricane trajectory prediction subdiscipline. Little research effort has been allocated to this particular problem space of high economic and social impact. This work is expected to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.

@@

Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}. As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape. A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP. A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km. The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudal points \cite{lat-long-distance}. The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.

@@

Assuming the input data is ordered chronogically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years. Recent trajectories will require less repitition to produce associaton rules, and higher repitition will be needed to generate frequent sequences from older data. The approach of chronogically weighted data will produce association rules that favor recent trends in hurricane trajectories. Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon. Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining. However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) weigh individual records based on a proportional scale and (3) allow records to have a weight less than 1.

>>>>

There are two categories of motivation behind this work. The first motivation of this work is to add to the total investigation completed in the data mining-based hurricane trajectory prediction sub-discipline. Little research effort has been allocated to this particular problem space of high economic and social impact. This work is expected to contribute to the growing effort by academic and government researchers to develop reliable methods of hurricane trajectory prediction in the interest of public safety and economic stability.

@@

Consider figure (2), generated using a sample of the recorded hurricane trajectories in HURDAT2 with the visualization tool Matplotlib \cite{matplotlib}. As shown in the pair of images, region discretization organizes continuous hurricane coordinate sequences into discrete region sequences of approximately uniform size and square shape. A discretization block size of 5 is selected in HTPDM and is replicated in WARD-HTP. A block size of 5 corresponds approximately to a real-world block dimension of 500km by 500km. The block dimension size is approximate because the mapping between coordinate degrees and kilometres is imprecise, and is dependent on location in the case of longitudinal points \cite{lat-long-distance}. The approximate size of 500km is similar to the average hurricane total diameter \cite{hurricane-distances}, permitting the coarse-grain expression of hurricane trajectories without undue loss in precision.

@@

Assuming the input data is ordered chronologically, introducing weighted support to AprioriAll will result in higher support for patterns which occur in recent years. Recent trajectories will require less repetition to produce association rules, and higher repetition will be needed to generate frequent sequences from older data. The approach of chronologically weighted data will produce association rules that favour recent trends in hurricane trajectories. Therefore, the introduction of weighted support to AprioriAll is expected to further increase the hurricane trajectory prediction correctness ratio for current and future trajectories being tested upon. Publications such as \cite{war-mining} and \cite{cai-weighted-thesis} first introduce the use of weighted support for association rule mining. However, to the knowledge of the authors this work is the first to (1) apply weighted support to sequential data, (2) weigh individual records based on a proportional scale and (3) allow records to have a weight less than 1.
